---
title: "The effect of dehorning and other potential factors influencing horn growth in game-ranched white rhinoceroses (Ceratotherium simum simum)"
output: github_document
---

This is a notebook for the dehorning data of Ververs et al. (2018). When you execute code within the notebook, the results appear beneath the code. 

## Statistical analysis preface

Currently the following R packages were loaded

```{r, echo=FALSE}
#read excel file
if (!require("openxlsx")) {
  install.packages("openxlsx", dependencies = TRUE)
  library(openxlsx)
}
#effects plot
if (!require("effects")) {
  install.packages("effects", dependencies = TRUE)
  library(effects)
}
# describe() function
if (!require("psych")) {
  install.packages("psych", dependencies = TRUE)
  library(psych)
}
# lsmeans() function
if (!require("lsmeans")) {
  install.packages("lsmeans", dependencies = TRUE)
  library(lsmeans)
}
# mfa() function
if (!require("FactoMineR")) {
  install.packages("FactoMineR", dependencies = TRUE)
  library(FactoMineR)
}
if (!require("factoextra")) {
  install.packages("factoextra", dependencies = TRUE)
  library(factoextra)
}
```

## Data extraction, transformation and loading

The data was read from the excel file containing all dehorings, and some 

```{r, echo=FALSE}
#Set correct path
path_csv <- "./dehorndata.csv"
#Read the data
df.cv <- read.csv2(path_csv)
#Create factors instead of characters
character_vars <- lapply(df.cv, class) == "character"
df.cv[, character_vars] <- lapply(df.cv[, character_vars], as.factor)
#Create extra date variables
#For birth date
df.cv$B_YEAR <- as.factor(df.cv$B_YEAR)
df.cv$B_MONTH <- as.factor(df.cv$B_MONTH)
df.cv$B_QUARTER <- as.factor(df.cv$B_QUARTER)
#And calving date
df.cv$C_YEAR <- as.factor(df.cv$C_YEAR)
df.cv$C_MONTH <- as.factor(df.cv$C_MONTH)
df.cv$C_QUARTER <- as.factor(df.cv$C_QUARTER)
#And dehorn date
df.cv$DH_YEAR <- as.factor(df.cv$DH_YEAR)
df.cv$DH_MONTH <- as.factor(df.cv$DH_MONTH)
df.cv$DH_QUARTER <- as.factor(df.cv$DH_QUARTER)

#Create numerical factor for number of dehornings
df.cv$N_DEHORNINGS <- as.factor(df.cv$N_DEHORNINGS)
#And a calculated parity (to avoid mistakes in parity number written down)
df.cv$CALC_PARITY <- as.factor(df.cv$CALC_PARITY)

#Our final dataset is called df.cv.final and contains specific columns
df.cv.final <- df.cv[, c("GENDER", 
                         "AH_W_G",
                         "PH_W_G",
                         "SHAVINGS_W_G",
                         "TOTAL", 
                         "AH_L",
                         "AH_C",
                         "AH_STUMP_L",
                         "PH_L",
                         "PH_C",
                         "PH_STUMP_L",
                         "CALC_AGE",
                         "CALC_ARRIVAL",
                         "DH_YEAR",
                         "DH_MONTH",
                         "DH_QUARTER",
                         "W_DIFF",
                         "L_DIFF",
                         "C_DIFF",
                         "B_YEAR",
                         "B_MONTH",
                         "B_QUARTER",
                         "DIM_DEHORN",
                         "INT_DEHORN",
                         "CALC_PARITY",
                         "C_YEAR",
                         "C_MONTH",
                         "C_QUARTER",
                         "N_DEHORNINGS",
                         "GR_DAYS",
                         "G_TOTAL",
                         "G_AH_W",
                         "G_AH_L",
                         "G_AH_C",
                         "G_PH_W",
                         "G_PH_L",
                         "G_PH_C")]
```

This has a first larger dataset is loaded in memory to start from.

# Descriptive statistics


A summary is requested holding [descriptive statistics]([https://en.wikipedia.org/wiki/Descriptive_statistics).

## All sexes together
```{r}
summary(df.cv.final)
describe(df.cv.final)
```

##Males

```{r}
summary(df.cv.final[df.cv.final$GENDER == "M", ])
describe(df.cv.final[df.cv.final$GENDER == "M", ])
```

##Females

```{r}
summary(df.cv.final[df.cv.final$GENDER == "F", ])
describe(df.cv.final[df.cv.final$GENDER == "F", ])
```

##Only first dehoring descriptives

```{r}
summary(df.cv.final[df.cv.final$N_DEHORNINGS == "1", ])
describe(df.cv.final[df.cv.final$N_DEHORNINGS == "1", ])
```

#Data analyis

Next, subsets are constructed for the separate hypothesis.

## Factors influencing the total growth of the horn in reproducing females

### Data preparation

We now have a subset of animals which hold the following information

```{r}
summary(na.omit(df.cv[, c("CALC_ARRIVAL", "DH_QUARTER", "DIM_DEHORN", "CALC_PARITY", "C_QUARTER", "N_DEHORNINGS", "G_TOTAL", "DAYSDUE")]))
#Histogram to explore days due
hist(df.cv[, c("DAYSDUE")],
     main = "Days due",
     xlab ="")
```

After data mining the aforementioned dataset, we decided to create a subset of COWS (number of dehornings larger than 1 with non-missing data for the following features:

1. Days since arrival at the farm at dehorning (proxy for age)
2. Quarter of dehorning (proxy for season)
3. Days since calving at dehorning
4. Calculated number of calvings
5. Quarter at calving
6. Number of dehornings
7. Total growth

```{r}
df.cv.cows <- na.omit(df.cv[!df.cv$N_DEHORNINGS == "1", c("CALC_ARRIVAL", "DH_QUARTER", "DIM_DEHORN", "CALC_PARITY", "C_QUARTER", "N_DEHORNINGS", "G_TOTAL")])
summary(df.cv.cows)
#Set all number of dehornings larger then 3 equal to 3 
df.cv.cows[df.cv.cows$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.cows[df.cv.cows$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
```

Some histograms to increase interpretation of the current dataset

```{r}
op = par(mfrow=c(1, 2))
hist(df.cv.cows$G_TOTAL,
     main = "Total growth (g/day)",
     xlab ="")
hist(log10(df.cv.cows$G_TOTAL),
     main = "Log10(Total growth (g/day))",
     xlab = "")
op = par(mfrow=c(1, 1))
hist(df.cv.cows$CALC_ARRIVAL,
     main = "Days since arrival at the farm",
     xlab ="")
op = par(mfrow=c(2, 2))
hist(df.cv.cows[df.cv.cows$N_DEHORNINGS == "2",]$CALC_ARRIVAL,
     main = "Days since arrival at the farm",
     xlab ="Dehorning 2")
hist(df.cv.cows[df.cv.cows$N_DEHORNINGS == "3",]$CALC_ARRIVAL,
     main = "Days since arrival at the farm",
     xlab ="Dehorning 3")
hist(log10(df.cv.cows[df.cv.cows$N_DEHORNINGS == "2",]$G_TOTAL),
     main = "Log10(Total growth (g/day))",
     xlab ="Dehorning 2")
hist(log10(df.cv.cows[df.cv.cows$N_DEHORNINGS == "3",]$G_TOTAL),
     main = "Log10(Total growth (g/day))",
     xlab ="Dehorning 3")

```

### Model building

A [generalised linear model](https://en.wikipedia.org/wiki/Generalized_linear_model)  is constructed, using [family="log"](http://stats.stackexchange.com/questions/47840/linear-model-with-log-transformed-response-vs-generalized-linear-model-with-log) function.

First the baseline model is build, and next [univariate models](https://en.wikipedia.org/wiki/Univariate) are constructed for each of the variables.

```{r, echo=FALSE}
## Creation of basemodel and univariates
summary(baseline <- glm(G_TOTAL~ 1, 
                        df.cv.cows, 
                        family = gaussian(link="log")))
summary(glm2 <- glm(G_TOTAL~ N_DEHORNINGS, 
                    df.cv.cows, 
                    family = gaussian(link="log")))
summary(glm3 <- glm(G_TOTAL~ CALC_ARRIVAL, 
                    df.cv.cows, 
                    family = gaussian(link="log")))
summary(glm4 <- glm(G_TOTAL~ CALC_PARITY, 
                    df.cv.cows, 
                    family = gaussian(link="log")))
summary(glm5 <- glm(G_TOTAL~ DIM_DEHORN, 
                    df.cv.cows, 
                    family = gaussian(link="log")))
summary(glm6 <- glm(G_TOTAL~ DH_QUARTER, 
                    df.cv.cows, 
                    family = gaussian(link="log")))
summary(glm7 <- glm(G_TOTAL~ C_QUARTER, 
                    df.cv.cows, 
                    family = gaussian(link="log")))
```

### Model comparison

```{r}
## Comparison of basemodel and single variate models
anova(baseline, glm2, glm3, glm4, glm5, glm6, glm7,  test = "Chisq")
```

We select the variables that significantly ameliorate the model fit.

```{r}
summary(glm8 <- glm(G_TOTAL~ N_DEHORNINGS + CALC_ARRIVAL + DH_QUARTER, 
                    df.cv.cows, 
                    family = gaussian(link="log")))
anova(baseline, glm8, test = "Chisq")
```

### Interaction

After exploration a one way interaction was found to significantly increase the model fit.

```{r}
summary(glm9 <- glm(G_TOTAL~ N_DEHORNINGS:CALC_ARRIVAL + DH_QUARTER, 
                    df.cv.cows, 
                    family = gaussian(link="log")))
anova(baseline, glm9, test = "Chisq")
```

To understand the effect of the interaction, an effects plot was created

```{r}
plot(effect("N_DEHORNINGS:CALC_ARRIVAL", 
            glm9, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
     main = "Interaction between number of dehornings and days at farm",
     xlab = "Days since arrival",
     ylab = "Total growth (g/day)")
```

### Final model least square means

The [least square means](https://en.wikipedia.org/wiki/Least_squares) are constructed for the final variables.

This is the overal P-value per variables

```{r, echo=FALSE}
anova(glm9, test = "Chisq")
```

Least square means, Tukey adjusted

```{r, echo=FALSE}

lsmeans(glm9, pairwise~N_DEHORNINGS, type = "response", adjust = "tukey" )
lsmeans(glm9, "CALC_ARRIVAL", type = "response")
lsmeans(glm9, pairwise~DH_QUARTER, type = "response", adjust = "tukey")
```

## Factors influencing the total growth of the horn in males and females

### Data preparation

From the previous model we learned that the moment of calving has no influence on the growth of the horn. Can we hence build a model accros sexes to increase the number of animals and get other insights.

```{r, echo=TRUE}
#Make sure to drop the extemely low growth which are caused by dirty data
df.cv.all <- na.omit(df.cv[df.cv$G_TOTAL < 100,
                           (names(df.cv) %in% c("CALC_AGE", 
                                                "DH_QUARTER", 
                                                "GENDER", 
                                                "G_TOTAL",
                                                "N_DEHORNINGS",
                                                "B_QUARTER"
                                                ))])
summary(df.cv.all)

df.cv.all[df.cv.all$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.all[df.cv.all$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"

```

### Univariate model building

```{r, echo=FALSE}
summary(baseline <- glm(G_TOTAL~ 1, 
                        df.cv.all, 
                        family = gaussian(link="log")))
summary(glm2 <- glm(G_TOTAL~ N_DEHORNINGS, 
                    df.cv.all, 
                    family = gaussian(link="log")))
summary(glm3 <- glm(G_TOTAL~ CALC_AGE, 
                    df.cv.all, 
                    family = gaussian(link="log")))
summary(glm4 <- glm(G_TOTAL~ DH_QUARTER, 
                    df.cv.all, 
                    family = gaussian(link="log")))
summary(glm5 <- glm(G_TOTAL~ B_QUARTER, 
                    df.cv.all, 
                    family = gaussian(link="log")))
summary(glm6 <- glm(G_TOTAL~ GENDER, 
                    df.cv.all, 
                    family = gaussian(link="log")))
```

### Model comparison

```{r, echo=FALSE}
## Comparison of basemodel and single variate models
anova(baseline, glm2, glm3, glm4, glm5, glm6, test = "Chisq")

anova(baseline, glm2, test = "Chisq")
anova(baseline, glm3, test = "Chisq")
anova(baseline, glm4, test = "Chisq")
anova(baseline, glm5, test = "Chisq")
anova(baseline, glm6, test = "Chisq")
```

### Model selection

```{r, echo=FALSE}
summary(glm7 <- glm(G_TOTAL~ N_DEHORNINGS + CALC_AGE + GENDER,
                    df.cv.all, 
                    family = gaussian(link="log")))
anova(glm7, test = "Chisq")
```

### Interaction

```{r}
summary(glm8 <- glm(G_TOTAL~ N_DEHORNINGS*CALC_AGE + GENDER, df.cv.all, family = gaussian(link="log")))
anova(glm8, test = "Chisq")
```
We conclude to drop the interaction in this model

### Final model least square means

```{r, echo=FALSE}
lsmeans(glm7, pairwise~N_DEHORNINGS, type = "response", adjust="tukey")
lsmeans(glm7, ~CALC_AGE, type = "response")
#lsmeans(glm7, ~N_DEHORNINGS*CALC_AGE, type = "response")
lsmeans(glm7, pairwise~ GENDER, type = "response")
```

## Hypothesis 3: Factors influencing the growth of the anterior horn weight in males and females

### Data preparation

```{r, echo=TRUE}
df.cv.all.f <- na.omit(df.cv[df.cv$G_AH_W < 100,
                             (names(df.cv) %in% c("CALC_AGE",
                                                  "DH_QUARTER", 
                                                  "GENDER", 
                                                  "G_AH_W",
                                                  "N_DEHORNINGS",
                                                  "B_QUARTER"))])
summary(df.cv.all.f)
df.cv.all.f[df.cv.all.f$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.all.f[df.cv.all.f$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
#create histogram of the front horn weiht
op = par(mfrow=c(1, 2))
hist((df.cv.all.f$G_AH_W), 
     main = "Anterior horn weight growth (g/day)",
     xlab ="")
hist(log10(df.cv.all.f$G_AH_W),
      main = "Log10",
     xlab ="")
```

### Model building

```{r, echo=FALSE}
summary(baseline <- glm(G_AH_W~ 1, 
                        df.cv.all.f, 
                        family = gaussian(link="log")))
summary(glm2 <- glm(G_AH_W~ N_DEHORNINGS, 
                    df.cv.all.f, 
                    family = gaussian(link="log")))
summary(glm3 <- glm(G_AH_W~ CALC_AGE, 
                    df.cv.all.f, 
                    family = gaussian(link="log")))
summary(glm4 <- glm(G_AH_W~ DH_QUARTER, 
                    df.cv.all.f, 
                    family = gaussian(link="log")))
summary(glm5 <- glm(G_AH_W~ B_QUARTER, 
                    df.cv.all.f, 
                    family = gaussian(link="log")))
summary(glm6 <- glm(G_AH_W~ GENDER, 
                    df.cv.all.f, 
                    family = gaussian(link="log")))
```

### Model selection

```{r, echo=FALSE}
anova(baseline, glm2, glm3, glm4, glm4b, glm5, glm6, test = "Chisq")
anova(baseline, glm2, test = "Chisq")
anova(baseline, glm3, test = "Chisq")
anova(baseline, glm4, test = "Chisq")
anova(baseline, glm5, test = "Chisq")
anova(baseline, glm6, test = "Chisq")
```

### Model selection

```{r, echo=FALSE}
summary(glm7 <- glm(G_AH_W~ N_DEHORNINGS + 
                      CALC_AGE + 
                      GENDER, df.cv.all.f, family = gaussian(link="log")))
anova(glm7, test = "Chisq")
```

### Interaction

```{r}
summary(glm8 <- glm(G_AH_W~ N_DEHORNINGS*CALC_AGE +
                      GENDER*CALC_AGE, df.cv.all.f, family = gaussian(link="log")))
anova(glm8, test = "Chisq")
plot(effect("N_DEHORNINGS:CALC_AGE", 
            glm8, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
     main= "Interaction between number of dehornings and age",
     ylab = "Anterior horn Weight growth (g/day)",
     xlab = "Age")
```

### Final model least square means

```{r, echo=FALSE}
lsmeans(glm8, pairwise~N_DEHORNINGS, type = "response", adjust="tukey")
lsmeans(glm8, ~CALC_AGE, type = "response")
lsmeans(glm8, ~N_DEHORNINGS*CALC_AGE, type = "response")
lsmeans(glm8, pairwise~ GENDER, type = "response")
```

## Hypothesis 4: Factors influencing the growth of the anterior horn length in males and females

### Data preparation

```{r, echo=TRUE}
df.cv.all.f.l <- na.omit(df.cv[df.cv$G_AH_L < 10,
                               (names(df.cv) %in% c("CALC_AGE", 
                                                    "DH_QUARTER", 
                                                    "GENDER", 
                                                    "G_AH_L",
                                                    "N_DEHORNINGS",
                                                    "B_QUARTER"
                                                    ))])
summary(df.cv.all.f.l)
df.cv.all.f.l[df.cv.all.f.l$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.all.f.l[df.cv.all.f.l$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
op = par(mfrow=c(1, 2))
hist(df.cv.all.f.l$G_AH_L,
     main = "Anterior horn length growth (mm/day)",
     xlab ="")
hist(log10(df.cv.all.f.l$G_AH_L),
     main = "LOG10",
     xlab ="")
```

### Model building

```{r, echo=FALSE}
summary(baseline <- glm(G_AH_L~ 1, df.cv.all.f.l, family = gaussian(link="log")))
summary(glm2 <- glm(G_AH_L~ N_DEHORNINGS, df.cv.all.f.l, family = gaussian(link="log")))
summary(glm3 <- glm(G_AH_L~ CALC_AGE, df.cv.all.f.l, family = gaussian(link="log")))
summary(glm4 <- glm(G_AH_L~ DH_QUARTER, df.cv.all.f.l, family = gaussian(link="log"))
summary(glm5 <- glm(G_AH_L~ B_QUARTER, df.cv.all.f.l, family = gaussian(link="log")))
summary(glm6 <- glm(G_AH_L~ GENDER, df.cv.all.f.l, family = gaussian(link="log")))
```

### Model selection

```{r, echo=FALSE}
anova(baseline, glm2, glm3, glm4, glm5, glm6, test = "Chisq")
anova(baseline, glm2, test = "Chisq")
anova(baseline, glm3, test = "Chisq")
anova(baseline, glm4, test = "Chisq")
anova(baseline, glm5, test = "Chisq")
anova(baseline, glm6, test = "Chisq")
```

### Final model selection

```{r, echo=FALSE}
summary(glm7 <- glm(G_AH_L~ N_DEHORNINGS + 
                            CALC_AGE + 
                            GENDER, df.cv.all.f.l, family = gaussian(link="log")))
anova(glm7, test = "Chisq")
```

### Interaction

```{r}
summary(glm8 <- glm(G_AH_L~ N_DEHORNINGS*CALC_AGE + 
                      GENDER, df.cv.all.f.l, family = gaussian(link="log")))
anova(glm8, test = "Chisq")
plot(effect("N_DEHORNINGS:CALC_AGE", 
            glm8, 
            multiline=TRUE))
```

### Final model least square means

```{r, echo=FALSE}
lsmeans(glm8, pairwise~N_DEHORNINGS, type = "response", adjust="tukey")
lsmeans(glm8, ~CALC_AGE, type = "response")
lsmeans(glm8, ~N_DEHORNINGS*CALC_AGE, type = "response")
lsmeans(glm8, pairwise~ GENDER, type = "response")
```

## Hypothesis 5: Factors influencing the growth of the anterior horn circumference (mm/day) in males and females 

### Data preparation

```{r, echo=TRUE}
df.cv.all.f.c <- na.omit(df.cv[df.cv$G_AH_C < 100,(names(df.cv) %in% c("CALC_AGE", 
                                                                     "DH_QUARTER", 
                                                                     "GENDER", 
                                                                     "G_AH_C",
                                                                     "N_DEHORNINGS",
                                                                     "B_QUARTER"
                                                                     ))])
df.cv.all.f.c[df.cv.all.f.c$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.all.f.c[df.cv.all.f.c$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
summary(df.cv.all.f.c)
op = par(mfrow=c(1, 2))
hist(df.cv.all.f.c$G_AH_C,
     main = "Anterior horn circumference growth(mm/day)",
     xlab ="")
hist(log10(df.cv.all.f.c$G_AH_C),
     main = "Log10",
     xlab ="")
```

### Model building

```{r, echo=FALSE}
summary(baseline <- glm(G_AH_C~ 1, df.cv.all.f.c, family = gaussian(link="log")))
summary(glm2 <- glm(G_AH_C~ N_DEHORNINGS, df.cv.all.f.c, family = gaussian(link="log")))
summary(glm3 <- glm(G_AH_C~ CALC_AGE, df.cv.all.f.c, family = gaussian(link="log")))
summary(glm4 <- glm(G_AH_C~ DH_QUARTER, df.cv.all.f.c, family = gaussian(link="log")))
summary(glm5 <- glm(G_AH_C~ B_QUARTER, df.cv.all.f.c, family = gaussian(link="log")))
summary(glm6 <- glm(G_AH_C~ GENDER, df.cv.all.f.c, family = gaussian(link="log")))
```

### Model selection

```{r, echo=FALSE}
anova(baseline, glm2, glm3, glm4, glm5, glm6, test = "Chisq")
anova(baseline, glm2, test = "Chisq")
anova(baseline, glm3, test = "Chisq")
anova(baseline, glm4, test = "Chisq")
anova(baseline, glm5, test = "Chisq")
anova(baseline, glm6, test = "Chisq")
```

### Final model selection

```{r, echo=FALSE}
summary(glm7 <- glm(G_AH_C~ N_DEHORNINGS + 
                      CALC_AGE + 
                      GENDER, df.cv.all.f.c, family = gaussian(link="log")))
anova(glm7, test = "Chisq")
```

### Interaction

```{r}
summary(glm8 <- glm(G_AH_C~ N_DEHORNINGS*CALC_AGE + GENDER
                      , df.cv.all.f.c, family = gaussian(link="log")))
anova(glm8, test = "Chisq")
plot(effect("N_DEHORNINGS:CALC_AGE", 
            glm8, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
            main= "Interaction between number of dehornings and age",
            ylab = "Circular growth of the front horn",
            xlab = "Age")
```

### Final model least square means

```{r, echo=FALSE}
lsmeans(glm8, pairwise~N_DEHORNINGS, type = "response", adjust="tukey")
lsmeans(glm8, ~CALC_AGE, type = "response")
lsmeans(glm8, ~N_DEHORNINGS*CALC_AGE, type = "response")
lsmeans(glm8, pairwise~ GENDER, type = "response")
```


# Multivariate analysis

### Data preparation

Data needs to be scaled and centered

```{r}
df.cv.mfa <- na.omit(df.cv[,c("GENDER",
                      "N_DEHORNINGS", 
                      "G_TOTAL",
                      "G_AH_W",
                      "G_AH_L",
                      "G_AH_C",
                      "CALC_AGE")])
df.cv.mfa$G_TOTAL <- scale(log10(df.cv.mfa$G_TOTAL), center = TRUE, scale = TRUE)
df.cv.mfa$G_AH_W <- scale(log10(df.cv.mfa$G_AH_W), center = TRUE, scale = TRUE)
df.cv.mfa$G_AH_L <- scale(log10(df.cv.mfa$G_AH_L), center = TRUE, scale = TRUE)
df.cv.mfa$G_AH_C <- scale(log10(df.cv.mfa$G_AH_C), center = TRUE, scale = TRUE)
df.cv.mfa$CALC_AGE <- scale(df.cv.mfa$CALC_AGE, center = TRUE, scale = TRUE)
df.cv.mfa[df.cv.mfa$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.mfa[df.cv.mfa$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
describe(df.cv.mfa)
```

### Factor analysis

```{r}
# www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/115-famd-factor-analysis-of-mixed-data-in-r-essentials/
cv.FAMD <- FAMD(df.cv.mfa)
```

Show the eigenvalues and variance and vizualize

```{r}
eig.val <- get_eigenvalue(cv.FAMD)
head(eig.val)
fviz_screeplot(cv.FAMD)
```

Plot the variables and overall contribution

```{r}
# Define the labels of tick marks
DIM1 <- c("Anterior horn growth",
          "Total growth",
          "Number of dehornings",
          "Anterior horn circumference",
          "Anterior horn length",
          "Age",
          "Sex")
fviz_famd_var(cv.FAMD, repel = TRUE)
# Contribution to the first dimension
fviz_contrib(cv.FAMD, "var", 
             axes = 1,
             x.lab = DIM1)
# Contribution to the second dimension
fviz_contrib(cv.FAMD, "var", 
             axes = 2)
```

Contribution of the quantitative variables on dimension 1 and 2

```{r}
quanti.var <- get_famd_var(cv.FAMD, "quanti.var")
quanti.var 
fviz_famd_var(cv.FAMD, "quanti.var", repel = TRUE,
              col.var = "black")
fviz_famd_var(cv.FAMD, "quanti.var", col.var = "contrib", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE)
```

Contribution of the qualitative variables

```{r}
fviz_famd_var(cv.FAMD, "quali.var", col.var = "contrib", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")
             )
```
             
Plot the individual observations

```{r}
flabels <- c("Number of dehornings", "Sex")

fviz_ellipses(cv.FAMD, c("N_DEHORNINGS", "GENDER"), 
              addEllipses = TRUE, 
              ellipse.type = "confidence",
              geom = "point",
              main = "Qualitative variables"
              )
```
             
