---
title: "Ververs et al. (2018) - Reviewers request"
output: 
  github_document:
    toc: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---

This is a notebook for the dehorning data of Ververs et al. (2018). This extra notebook shows the reviewers request to test random/repeated effects. 

## Statistical analysis preface

Currently the following R packages were loaded

```{r, echo=FALSE}
# lme4
if (!require("lme4")) {
  install.packages("lme4", dependencies = TRUE)
  library(lme4)
}
# effects
if (!require("effects")) {
  install.packages("effects", dependencies = TRUE)
  library(effects)
}
# car
if (!require("car")) {
  install.packages("car", dependencies = TRUE)
  library(car)
}
# sjPlot
if (!require("sjPlot")) {
  install.packages("sjPlot", dependencies = TRUE)
  library(sjPlot)
}
# lsmeans() function
if (!require("lsmeans")) {
  install.packages("lsmeans", dependencies = TRUE)
  library(lsmeans)
}
# EnvStats() function
if (!require("EnvStats")) {
  install.packages("EnvStats", dependencies = TRUE)
  library(EnvStats)
}
```

## Main data extraction, transformation and loading

The data was read from the excel file containing all dehornings, and some 

```{r, echo=FALSE}
#Set correct path
path_csv <- "./dehorndata.csv"
#Read the data
df.cv <- read.csv2(path_csv)
#Create factors instead of characters
character_vars <- lapply(df.cv, class) == "character"
df.cv[, character_vars] <- lapply(df.cv[, character_vars], as.factor)
#Create extra date variables
#For birth date
df.cv$B_YEAR <- as.factor(df.cv$B_YEAR)
df.cv$B_MONTH <- as.factor(df.cv$B_MONTH)
df.cv$B_QUARTER <- as.factor(df.cv$B_QUARTER)
#And calving date
df.cv$C_YEAR <- as.factor(df.cv$C_YEAR)
df.cv$C_MONTH <- as.factor(df.cv$C_MONTH)
df.cv$C_QUARTER <- as.factor(df.cv$C_QUARTER)
#And dehorn date
df.cv$DH_YEAR <- as.factor(df.cv$DH_YEAR)
df.cv$DH_MONTH <- as.factor(df.cv$DH_MONTH)
df.cv$DH_QUARTER <- as.factor(df.cv$DH_QUARTER)

#Create numerical factor for number of dehornings
df.cv$N_DEHORNINGS <- as.factor(df.cv$N_DEHORNINGS)
#And a calculated parity (to avoid mistakes in parity number written down)
df.cv$CALC_PARITY <- as.factor(df.cv$CALC_PARITY)

#Our final dataset is called df.cv.final and contains specific columns
df.cv.final <- df.cv[, c("ANIMAL",
                         "GENDER", 
                         "AH_W_G",
                         "PH_W_G",
                         "SHAVINGS_W_G",
                         "TOTAL", 
                         "AH_L",
                         "AH_C",
                         "AH_STUMP_L",
                         "PH_L",
                         "PH_C",
                         "PH_STUMP_L",
                         "CALC_AGE",
                         "CALC_ARRIVAL",
                         "DH_YEAR",
                         "DH_MONTH",
                         "DH_QUARTER",
                         "W_DIFF",
                         "L_DIFF",
                         "C_DIFF",
                         "B_YEAR",
                         "B_MONTH",
                         "B_QUARTER",
                         "DIM_DEHORN",
                         "INT_DEHORN",
                         "CALC_PARITY",
                         "C_YEAR",
                         "C_MONTH",
                         "C_QUARTER",
                         "N_DEHORNINGS",
                         "GR_DAYS",
                         "G_TOTAL",
                         "G_AH_W",
                         "G_AH_L",
                         "G_AH_C",
                         "G_PH_W",
                         "G_PH_L",
                         "G_PH_C")]
```

# Hypothesis 1 - Factors influencing the total growth of the horn in reproducing females

## Data preparation

```{r}
df.cv.cows <- na.omit(df.cv[!df.cv$N_DEHORNINGS == "1", c("ANIMAL","CALC_ARRIVAL", "DH_QUARTER", "DIM_DEHORN", "CALC_PARITY", "C_QUARTER", "N_DEHORNINGS", "G_TOTAL")])

#Set all number of dehornings larger then 3 equal to 3 
df.cv.cows[df.cv.cows$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.cows[df.cv.cows$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
#Not enough animals in the 4th parity
df.cv.cows[df.cv.cows$CALC_PARITY == "4",]$CALC_PARITY <- "3"
df.cv.cows$ANIMAL <- as.factor(df.cv.cows$ANIMAL)

summary(df.cv.cows)
xtabs(~DH_QUARTER+N_DEHORNINGS, df.cv.cows)
xtabs(~C_QUARTER+N_DEHORNINGS, df.cv.cows)
```

## Data transformations of continuous variables

```{r}
df.cv.cows$CALC_ARRIVAL_BC <- boxcoxTransform(df.cv.cows$CALC_ARRIVAL, lambda = 1)
df.cv.cows$CALC_ARRIVAL_STD <- scale(df.cv.cows$CALC_ARRIVAL)
hist(df.cv.cows$CALC_ARRIVAL)
hist(df.cv.cows$CALC_ARRIVAL_STD)
hist(df.cv.cows$CALC_ARRIVAL_BC)

df.cv.cows$DIM_DEHORN_BC <- boxcoxTransform(df.cv.cows$DIM_DEHORN, lambda = 0.5)
df.cv.cows$DIM_DEHORN_BC_STD <- scale(df.cv.cows$DIM_DEHORN_BC)
hist(df.cv.cows$DIM_DEHORN)
hist(df.cv.cows$DIM_DEHORN_BC)
hist(df.cv.cows$DIM_DEHORN_BC_STD)
```

## Model building

### Generalised linear model

For more info (here)[https://en.wikipedia.org/wiki/Generalized_linear_model]

```{r}
baselineGLM <- glm(G_TOTAL ~ 1, 
                         df.cv.cows, 
                         family = gaussian(link="log"))
```

### Generalized Linear Mixed-Effects Models

For more info [here](https://en.wikipedia.org/wiki/Generalized_linear_mixed_model)

We try to take into account the effect of the repeated measures (N_DEHORNINGS) within each animal (ANIMAL). We first want to understand 

```{r}
baselineGLMM <- glmer(G_TOTAL ~ 1 + (1 | ANIMAL), 
                  df.cv.cows, 
                  family = gaussian(link="log"),
                  control=glmerControl(optimizer="bobyqa")
                  )
```

### GLM vs GLMM model comparison

```{r}
anova(baselineGLMM, baselineGLM, test="Chisq")
```

It seems that the random effect GLMM is increasing the model fit.  Hence we will proceed.

### Univariate models


```{r, echo=FALSE}
summary(glmer1 <- glmer(G_TOTAL ~ N_DEHORNINGS + (1| ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")
                  ))
summary(glmer2 <- glmer(G_TOTAL ~ CALC_ARRIVAL_STD + (1| ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")
                ))
summary(glmer3 <- glmer(G_TOTAL ~ CALC_PARITY + (1| ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")
                  ))
summary(glmer4 <- glmer(G_TOTAL ~ DIM_DEHORN_BC_STD + (1| ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")
                  ))
summary(glmer5 <- glmer(G_TOTAL ~ DH_QUARTER + (1| ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")
                ))
summary(glmer6 <- glmer(G_TOTAL ~ C_QUARTER + (1| ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")
                ))

```

### Univariate model comparison

```{r}
## Comparison of basemodel and single variate models
anova(baselineGLMM, glmer1 ,glmer2, glmer3, glmer4, glmer5, glmer6,  test = "Chisq")
```

### Multivariate model building

```{r}
summary(glmer10 <- glmer(G_TOTAL ~ DH_QUARTER*DIM_DEHORN_BC_STD    
                         +  (1 | ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")))
Anova(glmer10)
```

```{r}
summary(glmer11 <- glmer(G_TOTAL ~ N_DEHORNINGS + DH_QUARTER*CALC_ARRIVAL_STD 
                         +  (1|ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")))
Anova(glmer11)
```

## Final model selection 

```{r}
anova(glmer10, glmer11, test = "Chisq")
```

### Interaction modelling

To understand the effect of the interaction, an effects plot was created

```{r}
plot(effect("DH_QUARTER:DIM_DEHORN_BC_STD", 
            glmer10, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
     main = "Interaction between dehorn season and days since last dehorn",
     xlab = "Days since arrival",
     ylab = "Total growth (g/day)")
```

### Least square means

The [least square means](https://en.wikipedia.org/wiki/Least_squares) are constructed for the final variables.

This is the overal P-value per variables

```{r, echo=FALSE}
Anova(glmer10)
```

Least square means, Tukey adjusted

```{r, echo=FALSE}

lsmeans(glmer10, pairwise~DH_QUARTER, type = "response", adjust = "tukey" )

if (!require("multcompView")) {
  install.packages("multcompView", dependencies = TRUE)
  library(multcompView)
}
## http://rcompanion.org/handbook/G_06.html
cld(lsmeans(glmer10, pairwise~DH_QUARTER),
          alpha=0.05,
          Letters=letters,
          adjust="tukey")
```

# Hypothesis 2 - Factors influencing the total growth of the horn in males and females

## Data preparation

```{r}
#Make sure to drop the extemely low growth which are caused by dirty data
df.cv.all <- na.omit(df.cv[df.cv$G_TOTAL < 100,
                           (names(df.cv) %in% c("ANIMAL",
                                                "CALC_AGE", 
                                                "DH_QUARTER", 
                                                "GENDER", 
                                                "G_TOTAL",
                                                "N_DEHORNINGS",
                                                "B_QUARTER"
                                                ))])
df.cv.all$ANIMAL <- as.factor(df.cv.all$ANIMAL)
df.cv.all[df.cv.all$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.all[df.cv.all$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"

summary(df.cv.all)

xtabs(~DH_QUARTER+N_DEHORNINGS, df.cv.all)
xtabs(~B_QUARTER+N_DEHORNINGS, df.cv.all)
```

## Data transformations of continuous variables

```{r}
df.cv.all$CALC_AGE_BC <- boxcoxTransform(df.cv.all$CALC_AGE, lambda = .25)
df.cv.all$CALC_AGE_BC_STD <- scale(df.cv.all$CALC_AGE_BC)
hist(df.cv.all$CALC_AGE)
hist(df.cv.all$CALC_AGE_BC_STD)

```


## Model building

### Generalaised linear model

For more info [here]([https://en.wikipedia.org/wiki/Generalized_linear_model)

```{r}
baselineGLM <- glm(G_TOTAL ~ 1, 
                         df.cv.all, 
                         family = gaussian(link="log"))
```

### Generalized Linear Mixed-Effects Models

For more info [here](https://en.wikipedia.org/wiki/Generalized_linear_mixed_model)

We try to take into account the effect of the repeated measures (N_DEHORNINGS) within each animal (ANIMAL). We first want to understand 

```{r}
baselineGLMM <- glmer(G_TOTAL ~ 1 + (1 | ANIMAL), 
                  df.cv.all, 
                  family = gaussian(link="log"),
                  control=glmerControl(optimizer="bobyqa")
                  )
```


### Model comparison

```{r}
anova(baselineGLMM, baselineGLM, test="Chisq")
```
The GLMM performs better


### Univariate model building

```{r, echo=FALSE}
glmer1 <- glmer(G_TOTAL~ N_DEHORNINGS + (1 | ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
glmer2 <- glmer(G_TOTAL~ CALC_AGE_BC_STD + (1 | ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
glmer3 <- glmer(G_TOTAL~ DH_QUARTER + (1 | ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
glmer4 <- glmer(G_TOTAL~ B_QUARTER + (1 | ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
glmer5 <- glmer(G_TOTAL~ GENDER + (1 | ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
```

### Model comparison

```{r}
## Comparison of basemodel and single variate models
anova(baselineGLMM, glmer1 ,glmer2, glmer3, glmer4, glmer5, test = "Chisq")
```

### Model building

```{r}
glmer10 <- glmer(G_TOTAL ~ CALC_AGE_BC_STD + N_DEHORNINGS + B_QUARTER    
                         +  (1| ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
summary(glmer10)
Anova(glmer10)
```

```{r}
glmer11 <- glmer(G_TOTAL ~ N_DEHORNINGS*CALC_AGE_BC_STD + B_QUARTER    
                         +  (1| ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
summary(glmer11)
Anova(glmer11)

```

## Final model 

Test interaction fit

```{r}
anova(glmer10, glmer11, test="Chisq")
```

The model with the interaction fits best, lowest AIC

### Interaction modelling

To understand the effect of the interaction, an effects plot was created

```{r}
plot(effect("N_DEHORNINGS:CALC_AGE_BC_STD", 
            glmer11, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
     main = "Interaction between number of dehornings and days at farm",
     xlab = "Days since arrival",
     ylab = "Total growth (g/day)")
```

Next piece is used to show it on non transformed scale

```{r}
glmer11b <- glmer(G_TOTAL ~ N_DEHORNINGS*CALC_AGE + B_QUARTER    
                         +  (1| ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
plot(effect("N_DEHORNINGS:CALC_AGE", 
            glmer11b, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
     main = "Interaction between number of dehornings and days at farm",
     xlab = "Days since arrival",
     ylab = "Total growth (g/day)")
```

### Least square means

The [least square means](https://en.wikipedia.org/wiki/Least_squares) are constructed for the final variables.

This is the overal P-value per variables

```{r, echo=FALSE}
Anova(glmer11)
```

Least square means, Tukey adjusted

```{r, echo=FALSE}

lsmeans(glmer11, pairwise~N_DEHORNINGS, type = "response", adjust = "tukey" )
lsmeans(glmer11, pairwise~B_QUARTER, type = "response", adjust = "tukey")
# If you need the subscripts which differ
cld(lsmeans(glmer11, pairwise~N_DEHORNINGS),
          alpha=0.05,
          Letters=letters,
          adjust="tukey")
cld(lsmeans(glmer11, pairwise~B_QUARTER),
          alpha=0.05,
          Letters=letters,
          adjust="tukey")
```

# Hypothesis 3: Factors influencing the growth of the anterior horn weight in males and females

## Data preparation

```{r, echo=TRUE}
df.cv.all.f <- na.omit(df.cv[df.cv$G_AH_W < 100,
                             (names(df.cv) %in% c("ANIMAL",
                                                  "CALC_AGE",
                                                  "DH_QUARTER", 
                                                  "GENDER", 
                                                  "G_AH_W",
                                                  "N_DEHORNINGS",
                                                  "B_QUARTER"))])
summary(df.cv.all.f)
df.cv.all.f[df.cv.all.f$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.all.f[df.cv.all.f$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
#create histogram of the front horn weiht
op = par(mfrow=c(1, 2))
hist((df.cv.all.f$G_AH_W), 
     main = "Anterior horn weight growth (g/day)",
     xlab ="")
hist(log10(df.cv.all.f$G_AH_W),
      main = "Log10",
     xlab ="")
```

## Data transformations of continuous variables

```{r}
df.cv.all.f$CALC_AGE_BC <- boxcoxTransform(df.cv.all.f$CALC_AGE, lambda = .25)
df.cv.all.f$CALC_AGE_BC_STD <- scale(df.cv.all.f$CALC_AGE_BC)
hist(df.cv.all.f$CALC_AGE)
hist(df.cv.all.f$CALC_AGE_BC_STD)
```

##  Model building

### Generalaised linear model

For more info [here]([https://en.wikipedia.org/wiki/Generalized_linear_model)

```{r}
baselineGLM <- glm(G_AH_W ~ 1, 
                         df.cv.all.f, 
                         family = gaussian(link="log"))
```

### Generalized Linear Mixed-Effects Models

For more info [here](https://en.wikipedia.org/wiki/Generalized_linear_mixed_model)

We try to take into account the effect of the repeated measures (N_DEHORNINGS) within each animal (ANIMAL). We first want to understand 

```{r}
baselineGLMM <- glmer(G_AH_W ~ 1 + (1 | ANIMAL), 
                  df.cv.all.f, 
                  family = gaussian(link="log"),
                  control=glmerControl(optimizer="bobyqa")
                  )
```


### Model comparison

```{r}
anova(baselineGLMM, baselineGLM, test="Chisq")
```
The GLMM performs better

### Univariate model building

### Model building

```{r, echo=FALSE}
summary(glmer1 <- glmer(G_AH_W~ N_DEHORNINGS + (1 | ANIMAL), 
                    df.cv.all.f, 
                    family = gaussian(link="log")))
summary(glmer2 <- glmer(G_AH_W~ CALC_AGE_BC_STD + (1 | ANIMAL), 
                    df.cv.all.f, 
                    family = gaussian(link="log")))
summary(glmer3 <- glmer(G_AH_W~ DH_QUARTER + (1 | ANIMAL), 
                    df.cv.all.f, 
                    family = gaussian(link="log")))
summary(glmer4 <- glmer(G_AH_W~ B_QUARTER + (1 | ANIMAL), 
                    df.cv.all.f, 
                    family = gaussian(link="log")))
summary(glmer5 <- glmer(G_AH_W~ GENDER + (1 | ANIMAL), 
                    df.cv.all.f, 
                    family = gaussian(link="log")))
```

### Model selection

```{r, echo=FALSE}
anova(baselineGLMM, glmer1, glmer2, glmer3, glmer4, glmer5, test = "Chisq")
```

### Model selection

```{r, echo=FALSE}
summary(glm7 <- glm(G_AH_W~ N_DEHORNINGS + 
                      CALC_AGE + 
                      GENDER, df.cv.all.f, family = gaussian(link="log")))
anova(glm7, test = "Chisq")
```

### Interaction

```{r}
summary(glm8 <- glm(G_AH_W~ N_DEHORNINGS*CALC_AGE +
                      GENDER*CALC_AGE, df.cv.all.f, family = gaussian(link="log")))
anova(glm8, test = "Chisq")
plot(effect("N_DEHORNINGS:CALC_AGE", 
            glm8, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
     main= "Interaction between number of dehornings and age",
     ylab = "Anterior horn Weight growth (g/day)",
     xlab = "Age")
```

### Final model least square means

```{r, echo=FALSE}
lsmeans(glm8, pairwise~N_DEHORNINGS, type = "response", adjust="tukey")
lsmeans(glm8, ~CALC_AGE, type = "response")
lsmeans(glm8, ~N_DEHORNINGS*CALC_AGE, type = "response")
lsmeans(glm8, pairwise~ GENDER, type = "response")
```

## Hypothesis 4: Factors influencing the growth of the anterior horn length in males and females

### Data preparation

```{r, echo=TRUE}
df.cv.all.f.l <- na.omit(df.cv[df.cv$G_AH_L < 10,
                               (names(df.cv) %in% c("CALC_AGE", 
                                                    "DH_QUARTER", 
                                                    "GENDER", 
                                                    "G_AH_L",
                                                    "N_DEHORNINGS",
                                                    "B_QUARTER"
                                                    ))])
df.cv.all.f.l[df.cv.all.f.l$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.all.f.l[df.cv.all.f.l$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
summary(df.cv.all.f.l)
op = par(mfrow=c(1, 2))
hist(df.cv.all.f.l$G_AH_L,
     main = "Anterior horn length growth (mm/day)",
     xlab ="")
hist(log10(df.cv.all.f.l$G_AH_L),
     main = "LOG10",
     xlab ="")
```

### Model building

```{r, echo=FALSE}
summary(baseline <- glm(G_AH_L~ 1, df.cv.all.f.l, family = gaussian(link="log")))
summary(glm2 <- glm(G_AH_L~ N_DEHORNINGS, df.cv.all.f.l, family = gaussian(link="log")))
summary(glm3 <- glm(G_AH_L~ CALC_AGE, df.cv.all.f.l, family = gaussian(link="log")))
summary(glm4 <- glm(G_AH_L~ DH_QUARTER, df.cv.all.f.l, family = gaussian(link="log")))
summary(glm5 <- glm(G_AH_L~ B_QUARTER, df.cv.all.f.l, family = gaussian(link="log")))
summary(glm6 <- glm(G_AH_L~ GENDER, df.cv.all.f.l, family = gaussian(link="log")))
```

### Model selection

```{r, echo=FALSE}
anova(baseline, glm2, glm3, glm4, glm5, glm6, test = "Chisq")
anova(baseline, glm2, test = "Chisq")
anova(baseline, glm3, test = "Chisq")
anova(baseline, glm4, test = "Chisq")
anova(baseline, glm5, test = "Chisq")
anova(baseline, glm6, test = "Chisq")
```

### Final model selection

```{r, echo=FALSE}
summary(glm7 <- glm(G_AH_L~ N_DEHORNINGS + 
                            CALC_AGE + 
                            GENDER, df.cv.all.f.l, family = gaussian(link="log")))
anova(glm7, test = "Chisq")
```

### Interaction

```{r}
summary(glm8 <- glm(G_AH_L~ N_DEHORNINGS*CALC_AGE + 
                      GENDER, df.cv.all.f.l, family = gaussian(link="log")))
anova(glm8, test = "Chisq")
plot(effect("N_DEHORNINGS:CALC_AGE", 
            glm8, 
            multiline=TRUE))
```

### Final model least square means

```{r, echo=FALSE}
lsmeans(glm8, pairwise~N_DEHORNINGS, type = "response", adjust="tukey")
lsmeans(glm8, ~CALC_AGE, type = "response")
lsmeans(glm8, ~N_DEHORNINGS*CALC_AGE, type = "response")
lsmeans(glm8, pairwise~ GENDER, type = "response")
```

## Hypothesis 5: Factors influencing the growth of the anterior horn circumference (mm/day) in males and females 

### Data preparation

```{r, echo=TRUE}
df.cv.all.f.c <- na.omit(df.cv[df.cv$G_AH_C < 100,(names(df.cv) %in% c("CALC_AGE", 
                                                                     "DH_QUARTER", 
                                                                     "GENDER", 
                                                                     "G_AH_C",
                                                                     "N_DEHORNINGS",
                                                                     "B_QUARTER"
                                                                     ))])
df.cv.all.f.c[df.cv.all.f.c$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.all.f.c[df.cv.all.f.c$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
summary(df.cv.all.f.c)
op = par(mfrow=c(1, 2))
hist(df.cv.all.f.c$G_AH_C,
     main = "Anterior horn circumference growth(mm/day)",
     xlab ="")
hist(log10(df.cv.all.f.c$G_AH_C),
     main = "Log10",
     xlab ="")
```

### Model building

```{r, echo=FALSE}
summary(baseline <- glm(G_AH_C~ 1, df.cv.all.f.c, family = gaussian(link="log")))
summary(glm2 <- glm(G_AH_C~ N_DEHORNINGS, df.cv.all.f.c, family = gaussian(link="log")))
summary(glm3 <- glm(G_AH_C~ CALC_AGE, df.cv.all.f.c, family = gaussian(link="log")))
summary(glm4 <- glm(G_AH_C~ DH_QUARTER, df.cv.all.f.c, family = gaussian(link="log")))
summary(glm5 <- glm(G_AH_C~ B_QUARTER, df.cv.all.f.c, family = gaussian(link="log")))
summary(glm6 <- glm(G_AH_C~ GENDER, df.cv.all.f.c, family = gaussian(link="log")))
```

### Model selection

```{r, echo=FALSE}
anova(baseline, glm2, glm3, glm4, glm5, glm6, test = "Chisq")
anova(baseline, glm2, test = "Chisq")
anova(baseline, glm3, test = "Chisq")
anova(baseline, glm4, test = "Chisq")
anova(baseline, glm5, test = "Chisq")
anova(baseline, glm6, test = "Chisq")
```

### Final model selection

```{r, echo=FALSE}
summary(glm7 <- glm(G_AH_C~ N_DEHORNINGS + 
                      CALC_AGE + 
                      GENDER, df.cv.all.f.c, family = gaussian(link="log")))
anova(glm7, test = "Chisq")
```

### Interaction

```{r}
summary(glm8 <- glm(G_AH_C~ N_DEHORNINGS*CALC_AGE + GENDER
                      , df.cv.all.f.c, family = gaussian(link="log")))
anova(glm8, test = "Chisq")
plot(effect("N_DEHORNINGS:CALC_AGE", 
            glm8, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
            main= "Interaction between number of dehornings and age",
            ylab = "Circular growth of the front horn",
            xlab = "Age")
```

### Final model least square means

```{r, echo=FALSE}
lsmeans(glm8, pairwise~N_DEHORNINGS, type = "response", adjust="tukey")
lsmeans(glm8, ~CALC_AGE, type = "response")
lsmeans(glm8, ~N_DEHORNINGS*CALC_AGE, type = "response")
lsmeans(glm8, pairwise~ GENDER, type = "response")
```


