---
title: "Ververs et al. (2018) - Reviewers request"
output: 
  github_document:
    toc: true
    toc_depth: 3
editor_options: 
  chunk_output_type: inline
---

This is a notebook for the dehorning data of Ververs et al. (2018). This extra notebook shows the reviewers request to test random/repeated effects. 

## Statistical analysis preface

Currently the following R packages were loaded

```{r, echo=FALSE}
# lme4
if (!require("lme4")) {
  install.packages("lme4", dependencies = TRUE)
  library(lme4)
}
# effects
if (!require("effects")) {
  install.packages("effects", dependencies = TRUE)
  library(effects)
}
# car
if (!require("car")) {
  install.packages("car", dependencies = TRUE)
  library(car)
}
if (!require("mvtnorm")) {
  install.packages("mvtnorm", dependencies = TRUE)
  library(mvtnorm)
}
# sjPlot
if (!require("sjPlot")) {
  install.packages("sjPlot", dependencies = TRUE)
  library(sjPlot)
}
# lsmeans() function
if (!require("lsmeans")) {
  install.packages("lsmeans", dependencies = TRUE)
  library(lsmeans)
}
# EnvStats() function
if (!require("EnvStats")) {
  install.packages("EnvStats", dependencies = TRUE)
  library(EnvStats)
}
```

## Main data extraction, transformation and loading

The data was read from the excel file containing all dehornings, and some 

```{r, echo=FALSE}
#Set correct path
path_csv <- "./dehorndata.csv"
#Read the data
df.cv <- read.csv2(path_csv)
#Create factors instead of characters
character_vars <- lapply(df.cv, class) == "character"
df.cv[, character_vars] <- lapply(df.cv[, character_vars], as.factor)
#Create extra date variables
#For birth date
df.cv$B_YEAR <- as.factor(df.cv$B_YEAR)
df.cv$B_MONTH <- as.factor(df.cv$B_MONTH)
df.cv$B_QUARTER <- as.factor(df.cv$B_QUARTER)
#And calving date
df.cv$C_YEAR <- as.factor(df.cv$C_YEAR)
df.cv$C_MONTH <- as.factor(df.cv$C_MONTH)
df.cv$C_QUARTER <- as.factor(df.cv$C_QUARTER)
#And dehorn date
df.cv$DH_YEAR <- as.factor(df.cv$DH_YEAR)
df.cv$DH_MONTH <- as.factor(df.cv$DH_MONTH)
df.cv$DH_QUARTER <- as.factor(df.cv$DH_QUARTER)

#Create numerical factor for number of dehornings
df.cv$N_DEHORNINGS <- as.factor(df.cv$N_DEHORNINGS)
#And a calculated parity (to avoid mistakes in parity number written down)
df.cv$CALC_PARITY <- as.factor(df.cv$CALC_PARITY)

#Our final dataset is called df.cv.final and contains specific columns
df.cv.final <- df.cv[, c("ANIMAL",
                         "GENDER", 
                         "AH_W_G",
                         "PH_W_G",
                         "SHAVINGS_W_G",
                         "TOTAL", 
                         "AH_L",
                         "AH_C",
                         "AH_STUMP_L",
                         "PH_L",
                         "PH_C",
                         "PH_STUMP_L",
                         "CALC_AGE",
                         "CALC_ARRIVAL",
                         "DH_YEAR",
                         "DH_MONTH",
                         "DH_QUARTER",
                         "W_DIFF",
                         "L_DIFF",
                         "C_DIFF",
                         "B_YEAR",
                         "B_MONTH",
                         "B_QUARTER",
                         "DIM_DEHORN",
                         "INT_DEHORN",
                         "CALC_PARITY",
                         "C_YEAR",
                         "C_MONTH",
                         "C_QUARTER",
                         "N_DEHORNINGS",
                         "GR_DAYS",
                         "G_TOTAL",
                         "G_AH_W",
                         "G_AH_L",
                         "G_AH_C",
                         "G_PH_W",
                         "G_PH_L",
                         "G_PH_C")]
```

# Hypothesis 1 - Factors influencing the total growth of the horn in reproducing females

## Data preparation

```{r}
df.cv.cows <- na.omit(df.cv[!df.cv$N_DEHORNINGS == "1", c("ANIMAL","CALC_ARRIVAL", "DH_QUARTER", "DIM_DEHORN", "CALC_PARITY", "C_QUARTER", "N_DEHORNINGS", "G_TOTAL")])

#Set all number of dehornings larger then 3 equal to 3 
df.cv.cows[df.cv.cows$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.cows[df.cv.cows$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
df.cv.cows$N_DEHORNINGS <-  droplevels(df.cv.cows$N_DEHORNINGS)

#Not enough animals in the 4th parity
df.cv.cows[df.cv.cows$CALC_PARITY == "4",]$CALC_PARITY <- "3"
df.cv.cows$CALC_PARITY <- droplevels(df.cv.cows$CALC_PARITY)



df.cv.cows$ANIMAL <- as.factor(df.cv.cows$ANIMAL)

summary(df.cv.cows)
xtabs(~DH_QUARTER+N_DEHORNINGS, df.cv.cows)
xtabs(~C_QUARTER+N_DEHORNINGS, df.cv.cows)
```

## Data transformations of continuous variables

```{r fig.height=5, fig.width=10}
df.cv.cows$CALC_ARRIVAL_BC <- boxcoxTransform(df.cv.cows$CALC_ARRIVAL, lambda = 1)
df.cv.cows$CALC_ARRIVAL_BC_STD <- scale(df.cv.cows$CALC_ARRIVAL_BC)
hist(df.cv.cows$CALC_ARRIVAL)
hist(df.cv.cows$CALC_ARRIVAL_BC)
hist(df.cv.cows$CALC_ARRIVAL_BC_STD)

df.cv.cows$DIM_DEHORN_BC <- boxcoxTransform(df.cv.cows$DIM_DEHORN, lambda = 0.5)
df.cv.cows$DIM_DEHORN_BC_STD <- scale(df.cv.cows$DIM_DEHORN_BC)
hist(df.cv.cows$DIM_DEHORN)
hist(df.cv.cows$DIM_DEHORN_BC)
hist(df.cv.cows$DIM_DEHORN_BC_STD)
```

A box-cox transformation was performed in order to make certain variables normally distributed. Otherwise models would not fit

## Model building

### Generalized linear model

For more info (here)[https://en.wikipedia.org/wiki/Generalized_linear_model]

```{r}
baselineGLM <- glm(G_TOTAL ~ 1, 
                         df.cv.cows, 
                         family = gaussian(link="log"))
```

This is the baseline model without random effect of the animal

### Generalized Linear Mixed-Effects Models

For more info [here](https://en.wikipedia.org/wiki/Generalized_linear_mixed_model)

We try to take into account the effect of the repeated measures (N_DEHORNINGS) within each animal (ANIMAL). We first want to understand 

```{r}
baselineGLMM <- glmer(G_TOTAL ~ 1 + (1 | ANIMAL), 
                  df.cv.cows, 
                  family = gaussian(link="log"),
                  control=glmerControl(optimizer="bobyqa")
                  )
```

This is the baseline model with random effect of the animal

### GLM vs GLMM model comparison

```{r}
anova(baselineGLMM, baselineGLM, test="Chisq")
```

It seems that the random effect GLMM is increasing the model fit.  Hence we will proceed.

### Univariate models

```{r, echo=FALSE}
glmer1 <- glmer(G_TOTAL ~ N_DEHORNINGS + (1| ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")
                  )
glmer2 <- glmer(G_TOTAL ~ CALC_ARRIVAL_BC_STD + (1| ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")
                )
glmer3 <- glmer(G_TOTAL ~ CALC_PARITY + (1| ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")
                  )
glmer4 <- glmer(G_TOTAL ~ DIM_DEHORN_BC_STD + (1| ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")
                  )
glmer5 <- glmer(G_TOTAL ~ DH_QUARTER + (1| ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")
                )
glmer6 <- glmer(G_TOTAL ~ C_QUARTER + (1| ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa")
                )

```

### Univariate model comparison

```{r}
## Comparison of basemodel and single variate models
anova(baselineGLMM, glmer1 ,glmer2, glmer3, glmer4, glmer5, glmer6,  test = "Chisq")
```

Only model glmer4 and glmer5 improve model fit, we proceed with those. For example in the text it could say "Days in milk after dehorning (P<2e-16) and the quarter of dehorning (P<2e-16) affected the total growth". Eventually we do not write this as we need to also explore a multivariate model which contains both variables.

### Multivariate model building

```{r}
glmer7 <- glmer(G_TOTAL ~ DH_QUARTER + DIM_DEHORN_BC_STD    
                         +  (1 | ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
```

However, the interaction between both variables should be explored.

```{r}
glmer10 <- glmer(G_TOTAL ~ DH_QUARTER*DIM_DEHORN_BC_STD    
                         +  (1 | ANIMAL), 
                df.cv.cows, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
anova(glmer7, glmer10, test = "Chisq")
```

## Final model selection 

We finally choose the model with interactions because of lowest AIC (752)

```{r}
Anova(glmer10, test = "Chisq")
```

In the text it will say that "the quarter of dehorning (P<2e-16)" had on overall effect on the total growth". The LSMmeans (see further) can be used to show how this effect was. The "Days in milk" had an effect on the total growth within the quarter of dehorning. This effect can be explored using the effects plot.

### Interaction modelling

To understand the effect of the interaction, an effects plot was created

```{r fig.height=5, fig.width=10}
plot(effect("DH_QUARTER:DIM_DEHORN_BC_STD", 
            glmer10, 
            multiline=TRUE),
     main = "Interaction between dehorn season and bc_std days since last calving",
     xlab = "Days since last calving",
     ylab = "Total growth (g/day)")
```

Next piece is used to show it on non transformed scale

```{r fig.height=5, fig.width=10}
glmer10b <- glmer(G_TOTAL~ 
                        DH_QUARTER*DIM_DEHORN + 
                        (1|ANIMAL), 
                      df.cv.cows,
                      family = gaussian(link="log"),
                      control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)),
                      nAGQ = 2
                      )
plot(effect("DH_QUARTER:DIM_DEHORN", 
            glmer10b, 
            multiline=TRUE),
     main = "Interaction between dehorn season and days since last calving",
     xlab = "Days since last calving",
     ylab = "Total growth (g/day)")
```

### Least square means

The [least square means](https://en.wikipedia.org/wiki/Least_squares) are constructed for the final variables.

This is the overal P-value per variables

```{r, echo=FALSE}
Anova(glmer10)
```

Least square means, Tukey adjusted

```{r, echo=FALSE}
lsmeans(glmer10b, pairwise~DH_QUARTER, type = "response", adjust = "tukey" )
```

# Hypothesis 2 - Factors influencing the total growth of the horn in males and females

## Data preparation

```{r}
#Make sure to drop the extemely low growth which are caused by dirty data
df.cv.all <- na.omit(df.cv[df.cv$G_TOTAL < 100,
                           (names(df.cv) %in% c("ANIMAL",
                                                "CALC_AGE", 
                                                "DH_QUARTER", 
                                                "GENDER", 
                                                "G_TOTAL",
                                                "N_DEHORNINGS",
                                                "B_QUARTER"
                                                ))])
df.cv.all$ANIMAL <- as.factor(df.cv.all$ANIMAL)
df.cv.all[df.cv.all$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.all[df.cv.all$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"

summary(df.cv.all)

xtabs(~DH_QUARTER+N_DEHORNINGS, df.cv.all)
xtabs(~B_QUARTER+N_DEHORNINGS, df.cv.all)
```

## Data transformations of continuous variables

```{r}
df.cv.all$CALC_AGE_BC <- boxcoxTransform(df.cv.all$CALC_AGE, lambda = .25)
df.cv.all$CALC_AGE_BC_STD <- scale(df.cv.all$CALC_AGE_BC)
hist(df.cv.all$CALC_AGE)
hist(df.cv.all$CALC_AGE_BC_STD)

```


## Model building

### Generalized linear model

For more info [here]([https://en.wikipedia.org/wiki/Generalized_linear_model)

```{r}
baselineGLM <- glm(G_TOTAL ~ 1, 
                         df.cv.all, 
                         family = gaussian(link="log"))
```

### Generalized Linear Mixed-Effects Models

For more info [here](https://en.wikipedia.org/wiki/Generalized_linear_mixed_model)

We try to take into account the effect of the repeated measures (N_DEHORNINGS) within each animal (ANIMAL). We first want to understand 

```{r}
baselineGLMM <- glmer(G_TOTAL ~ 1 + (1 | ANIMAL), 
                  df.cv.all, 
                  family = gaussian(link="log"),
                  control=glmerControl(optimizer="bobyqa")
                  )
```


### GLM vs GLMM model comparison

```{r}
anova(baselineGLMM, baselineGLM, test="Chisq")
```
The GLMM performs better

### Univariate models

```{r, echo=FALSE}
glmer1 <- glmer(G_TOTAL~ N_DEHORNINGS + (1 | ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
glmer2 <- glmer(G_TOTAL~ CALC_AGE_BC_STD + (1 | ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
glmer3 <- glmer(G_TOTAL~ DH_QUARTER + (1 | ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
glmer4 <- glmer(G_TOTAL~ B_QUARTER + (1 | ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
glmer5 <- glmer(G_TOTAL~ GENDER + (1 | ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
```

### Univariate model comparison

```{r}
## Comparison of basemodel and single variate models
anova(baselineGLMM, glmer1 ,glmer2, glmer3, glmer4, glmer5, test = "Chisq")
```

### Multivariate model building

```{r}
glmer10 <- glmer(G_TOTAL ~ CALC_AGE_BC_STD + N_DEHORNINGS + B_QUARTER    
                         +  (1| ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
summary(glmer10)
Anova(glmer10)
```

## Final model selection

### Interaction modelling

```{r}
glmer11 <- glmer(G_TOTAL ~ N_DEHORNINGS*CALC_AGE_BC_STD + B_QUARTER    
                         +  (1| ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
summary(glmer11)
Anova(glmer11)

```

Test interaction fit

```{r}
anova(glmer10, glmer11, test="Chisq")
```

The model with the interaction fits best, lowest AIC

To understand the effect of the interaction, an effects plot was created

```{r fig.height=5, fig.width=10}
plot(effect("N_DEHORNINGS:CALC_AGE_BC_STD", 
            glmer11, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
     main = "Interaction between number of dehornings and bc_std days at farm",
     xlab = "Days since arrival",
     ylab = "Total growth (g/day)")
```

Next piece is used to show it on non transformed scale

```{r fig.height=5, fig.width=10}
glmer11b <- glmer(G_TOTAL ~ N_DEHORNINGS*CALC_AGE + B_QUARTER    
                         +  (1| ANIMAL), 
                df.cv.all, 
                family = gaussian(link="log"),
                control=glmerControl(optimizer="bobyqa"))
plot(effect("N_DEHORNINGS:CALC_AGE", 
            glmer11b, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
     main = "Interaction between number of dehornings and days at farm",
     xlab = "Days since arrival",
     ylab = "Total growth (g/day)")
```


### Least square means

The [least square means](https://en.wikipedia.org/wiki/Least_squares) are constructed for the final variables.

This is the overal P-value per variables

```{r, echo=FALSE}
Anova(glmer11b)
```

Least square means, Tukey adjusted

```{r, echo=FALSE}

lsmeans(glmer11b, pairwise~N_DEHORNINGS, type = "response", adjust = "tukey" )
lsmeans(glmer11b, pairwise~B_QUARTER, type = "response", adjust = "tukey")

```
### Interaction trends

```{r, echo=FALSE}
summary(glmer11b)
emtrends(glmer11b, pairwise ~ N_DEHORNINGS, var = "CALC_AGE")
```

# Hypothesis 3 - Factors influencing the growth of the anterior horn weight in males and females

## Data preparation

```{r fig.height=5, fig.width=10, echo=TRUE}
df.cv.all.f <- na.omit(df.cv[df.cv$G_AH_W < 100,
                             (names(df.cv) %in% c("ANIMAL",
                                                  "CALC_AGE",
                                                  "DH_QUARTER", 
                                                  "GENDER", 
                                                  "G_AH_W",
                                                  "N_DEHORNINGS",
                                                  "B_QUARTER"))])
summary(df.cv.all.f)
df.cv.all.f[df.cv.all.f$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.all.f[df.cv.all.f$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
#create histogram of the front horn weiht
op = par(mfrow=c(1, 2))
hist((df.cv.all.f$G_AH_W), 
     main = "Anterior horn weight growth (g/day)",
     xlab ="")
hist(log10(df.cv.all.f$G_AH_W),
      main = "Log10",
     xlab ="")
```

## Data transformations of continuous variables

```{r fig.height=5, fig.width=10}
df.cv.all.f$CALC_AGE_BC <- boxcoxTransform(df.cv.all.f$CALC_AGE, lambda = .25)
df.cv.all.f$CALC_AGE_BC_STD <- scale(df.cv.all.f$CALC_AGE_BC)
hist(df.cv.all.f$CALC_AGE)
hist(df.cv.all.f$CALC_AGE_BC_STD)
```

##  Model building

### Generalized linear model

For more info [here]([https://en.wikipedia.org/wiki/Generalized_linear_model)

```{r}
baselineGLM <- glm(G_AH_W ~ 1, 
                         df.cv.all.f, 
                         family = gaussian(link="log"))
```

### Generalized Linear Mixed-Effects Models

For more info [here](https://en.wikipedia.org/wiki/Generalized_linear_mixed_model)

We try to take into account the effect of the repeated measures (N_DEHORNINGS) within each animal (ANIMAL). We first want to understand 

```{r}
baselineGLMM <- glmer(G_AH_W ~ 1 + (1 | ANIMAL), 
                  df.cv.all.f, 
                  family = gaussian(link="log"),
                  control=glmerControl(optimizer="bobyqa")
                  )
```

### GLMM vs GLMM model comparison

```{r}
anova(baselineGLMM, baselineGLM, test="Chisq")
```
The GLMM performs better

### Univariate models

```{r, echo=FALSE}
glmer1 <- glmer(G_AH_W~ N_DEHORNINGS + (1 | ANIMAL), 
                    df.cv.all.f, 
                    family = gaussian(link="log"))
glmer2 <- glmer(G_AH_W~ CALC_AGE_BC_STD + (1 | ANIMAL), 
                    df.cv.all.f, 
                    family = gaussian(link="log"))
glmer3 <- glmer(G_AH_W~ DH_QUARTER + (1 | ANIMAL), 
                    df.cv.all.f, 
                    family = gaussian(link="log"))
glmer4 <- glmer(G_AH_W~ B_QUARTER + (1 | ANIMAL), 
                    df.cv.all.f, 
                    family = gaussian(link="log"))
glmer5 <- glmer(G_AH_W~ GENDER + (1 | ANIMAL), 
                    df.cv.all.f, 
                    family = gaussian(link="log"))
```

### Univariate model comparison

```{r, echo=FALSE}
anova(baselineGLMM, glmer1, glmer2, glmer3, glmer4, glmer5, test = "Chisq")
```

### Multivariate model building

```{r, echo=FALSE}
summary(glmer6 <- glmer(G_AH_W~ N_DEHORNINGS + 
                      CALC_AGE_BC_STD + 
                      B_QUARTER + 
                        (1 | ANIMAL), 
                      df.cv.all.f, family = gaussian(link="log")))
Anova(glmer6, test = "Chisq")
```

## Final model selection

### Interaction modelling

```{r fig.height=5, fig.width=10}
summary(glmer7 <- glmer(G_AH_W ~ 
                          N_DEHORNINGS*CALC_AGE_BC_STD +
                          B_QUARTER +
                          (1|ANIMAL), 
                        df.cv.all.f, 
                        family = gaussian(link="log")))
Anova(glmer7, test = "Chisq")
plot(effect("N_DEHORNINGS:CALC_AGE_BC_STD", 
            glmer7, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
     main= "Interaction between number of dehornings and bc_std age",
     ylab = "Anterior horn Weight growth (g/day)",
     xlab = "Age")
```

Next piece is used to show it on non transformed scale

```{r fig.height=5, fig.width=10}
glmer7b <- glmer(G_AH_W~ 
                        N_DEHORNINGS*CALC_AGE + 
                        B_QUARTER + 
                        (1|ANIMAL), 
                      df.cv.all.f, 
                      family = gaussian(link="log"),
                      control=glmerControl(optimizer="bobyqa")
                      )
plot(effect("N_DEHORNINGS:CALC_AGE", 
            glmer7b, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
     main= "Interaction between number of dehornings and age",
     ylab = "Anterior horn Weight growth (g/day)",
     xlab = "Age")
```

### Least square means

```{r, echo=FALSE}
lsmeans(glmer7b, pairwise~N_DEHORNINGS, type = "response", adjust="tukey")
lsmeans(glmer7b, pairwise~ B_QUARTER, type = "response")
```

### Interaction estimates

```{r, echo=FALSE}
summary(glmer7b)
emtrends(glmer7b, pairwise ~ N_DEHORNINGS, var = "CALC_AGE")
```

# Hypothesis 4 - Factors influencing the growth of the anterior horn length in males and females

## Data preparation

```{r fig.height=5, fig.width=10, echo=TRUE}
df.cv.all.f.l <- na.omit(df.cv[df.cv$G_AH_L < 10,
                               (names(df.cv) %in% c("ANIMAL",
                                                    "CALC_AGE", 
                                                    "DH_QUARTER", 
                                                    "GENDER", 
                                                    "G_AH_L",
                                                    "N_DEHORNINGS",
                                                    "B_QUARTER"
                                                    ))])
df.cv.all.f.l[df.cv.all.f.l$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.all.f.l[df.cv.all.f.l$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
summary(df.cv.all.f.l)
op = par(mfrow=c(1, 2))
hist(df.cv.all.f.l$G_AH_L,
     main = "Anterior horn length growth (mm/day)",
     xlab ="")
hist(log10(df.cv.all.f.l$G_AH_L),
     main = "LOG10",
     xlab ="")
```

## Data transformations of continuous variables

```{r fig.height=5, fig.width=10}
df.cv.all.f.l$CALC_AGE_BC <- boxcoxTransform(df.cv.all.f.l$CALC_AGE, lambda = .25)
df.cv.all.f.l$CALC_AGE_BC_STD <- scale(df.cv.all.f.l$CALC_AGE_BC)
hist(df.cv.all.f.l$CALC_AGE)
hist(df.cv.all.f.l$CALC_AGE_BC_STD)
```

## Model building

### Generalized linear model

For more info [here]([https://en.wikipedia.org/wiki/Generalized_linear_model)

```{r}
baselineGLM <- glm(G_AH_L ~ 1, 
                         df.cv.all.f.l, 
                         family = gaussian(link="log"))
```

### Generalized Linear Mixed-Effects Models

For more info [here](https://en.wikipedia.org/wiki/Generalized_linear_mixed_model)

We try to take into account the effect of the repeated measures (N_DEHORNINGS) within each animal (ANIMAL). We first want to understand 

```{r}
baselineGLMM <- glmer(G_AH_L ~ 1 + (1 | ANIMAL), 
                  df.cv.all.f.l, 
                  family = gaussian(link="log"),
                  control=glmerControl(optimizer="bobyqa")
                  )
```

### GLMM vs GLMM model comparison

```{r}
anova(baselineGLMM, baselineGLM, test="Chisq")
```
The GLMM performs better

### Univariate models


```{r, echo=FALSE}
glmer1 <- glmer(G_AH_L~ N_DEHORNINGS + (1|ANIMAL), 
                        df.cv.all.f.l, 
                        family = gaussian(link="log"),
                        control=glmerControl(optimizer="bobyqa")
                        )
glmer2 <- glmer(G_AH_L~ CALC_AGE_BC_STD + (1|ANIMAL), 
                        df.cv.all.f.l, 
                        family = gaussian(link="log"),
                        control=glmerControl(optimizer="bobyqa")
                        )
glmer3 <- glmer(G_AH_L~ DH_QUARTER + (1|ANIMAL), 
                        df.cv.all.f.l, 
                        family = gaussian(link="log"),
                        control=glmerControl(optimizer="bobyqa")
                        )
glmer4 <- glmer(G_AH_L~ B_QUARTER + (1|ANIMAL), 
                        df.cv.all.f.l, 
                        family = gaussian(link="log"),
                        control=glmerControl(optimizer="bobyqa")
                        )
glmer5 <- glmer(G_AH_L~ GENDER + (1|ANIMAL), 
                        df.cv.all.f.l, 
                        family = gaussian(link="log"),
                        control=glmerControl(optimizer="bobyqa")
                        )
```

### Univariate model comparison

```{r, echo=FALSE}
anova(baselineGLMM, glmer1, glmer2, glmer3, glmer4, glmer5, test = "Chisq")
```

### Multivariate model building

```{r, echo=FALSE}
summary(glmer6 <- glmer(G_AH_L~ N_DEHORNINGS + 
                            CALC_AGE_BC_STD + 
                            B_QUARTER + 
                          (1|ANIMAL), 
                        df.cv.all.f.l, 
                        family = gaussian(link="log"),
                        control=glmerControl(optimizer="bobyqa")
                        ))
Anova(glmer6, test = "Chisq")
```

## Final model selection

### Interaction modelling

```{r fig.height=5, fig.width=10}
summary(glmer8 <- glmer(G_AH_L~ 
                        N_DEHORNINGS*CALC_AGE_BC_STD + 
                        (1|ANIMAL), 
                      df.cv.all.f.l, 
                      family = gaussian(link="log"),
                      control=glmerControl(optimizer="bobyqa")
                      ))
Anova(glmer8, test = "Chisq")
plot(effect("N_DEHORNINGS:CALC_AGE_BC_STD", 
            glmer8, 
            multiline=TRUE),
     main = "Interaction between number of dehornings and bc_std days at farm",
     xlab = "Days at farm",
     ylab = "Total growth in length (g/day)")
```

Next piece is used to show it on non transformed scale

```{r fig.height=5, fig.width=10}
glmer8b <- glmer(G_AH_L~ 
                        N_DEHORNINGS*CALC_AGE + 
                        (1|ANIMAL), 
                      df.cv.all.f.l, 
                      family = gaussian(link="log"),
                      control=glmerControl(optimizer="bobyqa")
                      )
plot(effect("N_DEHORNINGS:CALC_AGE", 
            glmer8b, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
     main = "Interaction between number of dehornings and days at farm",
     xlab = "Days at farm",
     ylab = "Total growth in length (g/day)")
```


### Least square means

```{r, echo=FALSE}
lsmeans(glmer8b, ~N_DEHORNINGS|CALC_AGE, type = "response", adjust="tukey")
```

### Interaction estimates

```{r, echo=FALSE}
summary(glmer8b)
emtrends(glmer8b, pairwise ~ N_DEHORNINGS, var = "CALC_AGE")
```


# Hypothesis 5 - Factors influencing the growth of the anterior horn circumference (mm/day) in males and females 

## Data preparation

```{r fig.height=5, fig.width=10, echo=TRUE}
df.cv.all.f.c <- na.omit(df.cv[df.cv$G_AH_C < 100,(names(df.cv) %in% c(
                                                                     "ANIMAL",
                                                                     "CALC_AGE", 
                                                                     "DH_QUARTER", 
                                                                     "GENDER", 
                                                                     "G_AH_C",
                                                                     "N_DEHORNINGS",
                                                                     "B_QUARTER"
                                                                     ))])
df.cv.all.f.c[df.cv.all.f.c$N_DEHORNINGS == "4",]$N_DEHORNINGS <- "3"
df.cv.all.f.c[df.cv.all.f.c$N_DEHORNINGS == "5",]$N_DEHORNINGS <- "3"
summary(df.cv.all.f.c)
op = par(mfrow=c(1, 2))
hist(df.cv.all.f.c$G_AH_C,
     main = "Anterior horn circumference growth(mm/day)",
     xlab ="")
hist(log10(df.cv.all.f.c$G_AH_C),
     main = "Log10",
     xlab ="")
```

## Data transformations of continuous variables

```{r}
df.cv.all.f.c$CALC_AGE_BC <- boxcoxTransform(df.cv.all.f.c$CALC_AGE, lambda = .25)
df.cv.all.f.c$CALC_AGE_BC_STD <- scale(df.cv.all.f.c$CALC_AGE_BC)
hist(df.cv.all.f.c$CALC_AGE)
hist(df.cv.all.f.c$CALC_AGE_BC_STD)
```

## Model building

### Generalized linear model

For more info [here]([https://en.wikipedia.org/wiki/Generalized_linear_model)

```{r}
baselineGLM <- glm(G_AH_C ~ 1, 
                         df.cv.all.f.c, 
                         family = gaussian(link="log"))
```

### Generalized Linear Mixed-Effects Models

For more info [here](https://en.wikipedia.org/wiki/Generalized_linear_mixed_model)

We try to take into account the effect of the repeated measures (N_DEHORNINGS) within each animal (ANIMAL). We first want to understand 

```{r}
baselineGLMM <- glmer(G_AH_C ~ 1 + (1 | ANIMAL), 
                  df.cv.all.f.c, 
                  family = gaussian(link="log"),
                  control=glmerControl(optimizer="bobyqa")
                  )
```

### GLMM vs GLMM model comparison

```{r}
anova(baselineGLMM, baselineGLM, test="Chisq")
```
The GLMM performs better

### Univariate models

```{r, echo=FALSE}
glmer1 <- glmer(G_AH_C~ 
                          N_DEHORNINGS + (1|ANIMAL), 
                        df.cv.all.f.c, 
                        family = gaussian(link="log"),
                        control=glmerControl(optimizer="bobyqa")
                        )
glmer2 <- glmer(G_AH_C~ 
                          CALC_AGE_BC_STD + (1|ANIMAL), 
                        df.cv.all.f.c, 
                        family = gaussian(link="log"),
                        control=glmerControl(optimizer="bobyqa")
                        )
glmer3 <- glmer(G_AH_C~ 
                          DH_QUARTER + (1|ANIMAL), 
                        df.cv.all.f.c, 
                        family = gaussian(link="log"),
                        control=glmerControl(optimizer="bobyqa")
                        )
glmer4 <- glmer(G_AH_C~ 
                          B_QUARTER + (1|ANIMAL), 
                        df.cv.all.f.c, 
                        family = gaussian(link="log"),
                        control=glmerControl(optimizer="bobyqa")
                        )
glmer5 <- glmer(G_AH_C~ 
                          GENDER + (1|ANIMAL), 
                        df.cv.all.f.c, 
                        family = gaussian(link="log"),
                        control=glmerControl(optimizer="bobyqa")
                        )
```

### Univariate model comparison

```{r, echo=FALSE}
anova(baselineGLMM, glmer1, glmer2, glmer3, glmer4, glmer5, test = "Chisq")
```

### Multivariate model building

```{r, echo=FALSE}
summary(glmer7 <- glmer(G_AH_C~ 
                          N_DEHORNINGS + 
                          CALC_AGE_BC_STD + 
                          (1|ANIMAL), 
                      df.cv.all.f.c, 
                      family = gaussian(link="log"),
                      control=glmerControl(optimizer="bobyqa")
                      ))
Anova(glmer7, test = "Chisq")
```

## Final model selection

### Interaction modelling

```{r fig.height=5, fig.width=10}
summary(glmer8 <- glmer(G_AH_C~ 
                          N_DEHORNINGS*CALC_AGE_BC_STD + 
                          (1|ANIMAL), 
                        df.cv.all.f.c, 
                        family = gaussian(link="log"),
                        control=glmerControl(optimizer="bobyqa")
                        ))
anova(glmer8, test = "Chisq")
plot(effect("N_DEHORNINGS:CALC_AGE_BC_STD", 
            glmer8, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
            main= "Interaction between number of dehornings and bc_std age",
            ylab = "Circular growth of the front horn",
            xlab = "Age")
```

Next piece is used to show it on non transformed scale

```{r fig.height=5, fig.width=10}
glmer8b <- glmer(G_AH_C~ 
                        N_DEHORNINGS*CALC_AGE + 
                        B_QUARTER + 
                        (1|ANIMAL), 
                      df.cv.all.f.c, 
                      family = gaussian(link="log"),
                      control=glmerControl(optimizer="bobyqa")
                      )
plot(effect("N_DEHORNINGS:CALC_AGE", 
            glmer8b, 
            #xlevels=list(CALC_ARRIVAL=900:2000), 
            multiline=TRUE),
            main= "Interaction between number of dehornings and age",
            ylab = "Circular growth of the front horn",
            xlab = "Age")
```

### Least square means

```{r, echo=FALSE}
lsmeans(glmer8b, pairwise~N_DEHORNINGS, type = "response", adjust="tukey")

```

### Interaction estimates

```{r, echo=FALSE}
summary(glmer8b)
emtrends(glmer8b, pairwise ~ N_DEHORNINGS, var = "CALC_AGE")
```

